---
# No specific front matter needed unless customizing
---

# Exploratory Data Analysis (EDA)

```{r, "EDA"}
library(ggplot2)
library(forecast)

ts_data <- AirPassengers

# The data is a monthly time series of international airline passengers from 1949 to 1960.
# Let's explore basic properties.

# Convert to a standard ts object (it already is, but just reaffirming)
ts_data <- as.ts(ts_data)

start_year <- start(ts_data)[1]
end_year   <- end(ts_data)[1]
n_obs      <- length(ts_data)

cat("This dataset starts in", start_year, "and ends in", end_year, 
    "with", n_obs, "monthly observations.\n")

# Basic summary
summary(ts_data)
```

## Raw Time Series Plot

```{r, "Plot Raw Time Series Plot"}
# Create a data frame for ggplot
df <- data.frame(
  date = seq(as.Date("1949-01-01"), as.Date("1960-12-01"), by = "month"),
  passengers = as.numeric(ts_data)
)

# Plot the time series using ggplot2
ggplot(df, aes(x = date, y = passengers)) +
  geom_line(color = "blue") +
  labs(
    title = "AirPassengers: Monthly Totals (1949-1960)",
    x = "Year",
    y = "Passengers (1000s)"
  ) +
  theme_minimal()
```

## Time Series Decomposition

```{r, "Time Series Decomposition"}
# Decompose the time series using classical multiplicative decomposition
decomp_data <- decompose(ts_data, type = "multiplicative")

# Plot the decomposition using ggplot2 (via autoplot from the forecast package)
autoplot(decomp_data) +
  ggtitle("Decomposition of AirPassengers Time Series") +
  theme_minimal()
```

Observations from EDA:
- The number of passengers generally increases over time, indicating a growing trend.
- There is a clear seasonal pattern (peaks in mid-year).
- The variance grows over time, hinting at non-stationarity.
- The multiplicative decomposition visually shows strong seasonality and an upward trend.