---
# No specific front matter needed unless customizing
---

## Model Overview

TBATS (Trigonometric, Box-Cox transformation, ARMA errors, Trend, and Seasonal components) is designed to handle complex seasonal patterns and non-linear trends. It is especially useful for data with multiple seasonal periods or subtle non-linear behaviors.

## Model Fitting

For this analysis, we use the built-in `co2` dataset, which records atmospheric CO2 concentrations. We fit a TBATS model using the `tbats` function from the `forecast` package.

```{r, "Fit TBATS"}
library(forecast)
library(ggplot2)

ts_data <- AirPassengers
fit_tbats <- tbats(ts_data)
```

## Model Evaluation

We evaluate the TBATS model by forecasting the next 12 months and plotting the forecast. This visualization shows how well the model captures the underlying patterns in the `AirPassengers` data.

```{r, "Forecast TBATS"}
fc_tbats <- forecast(fit_tbats, h = 12)
autoplot(fc_tbats) + ggtitle("TBATS 12-Month Forecast")
```

### RMSE and MAPE Metrics

We calculate the Root Mean Squared Error (RMSE) and Mean Absolute Percentage Error (MAPE) using the accuracy function to evaluate the model’s performance on the training set.

```{r, "TBATS RMSE"}
acc_tbats <- accuracy(fit_tbats)
acc_tbats_table <- data.frame(
  Metric = c("RMSE", "MAPE"),
  Value = c(acc_tbats[1, "RMSE"], acc_tbats[1, "MAPE"])
)
knitr::kable(acc_tbats_table, caption = "TBATS Model Performance Metrics")
```

Evaluation Summary:
- TBATS summary provides insights into the transformation and seasonal components used.
- Forecast visualization aids in assessing the model’s predictive performance.